### 2025-06-14 Code Repository.Rmd
Some adjustments

---
title: "Code Repository"
author: "Brendan Woods"
date: "`r Sys.Date()`"
output: html_document
---

## Contents
- Pre-processing
- Creating the Composites using original EMERGE data
- Permutation Code
- Theoretical Distributions
- Latent Variable Analysis McMenamin et al.
- Power Estimation
- Type I error rate Estimation
- Simulating toy data set

## Pre-processing
```{r}
library(dplyr)
library(MASS)
library(tidyr)
library(mvtnorm)
library(ggplot2)
library(kableExtra)

load("\\Users\\brend\\Downloads\\data")
tmpDat <- tmpDat[!(is.na(tmpDat$InsulinYN) & is.na(tmpDat$gw32) & is.na(tmpDat$gw38)), ] ## removes individuals with no observations for 3 outcomes
tmpDat$binary_insulin<-0
tmpDat$binary_insulin[tmpDat$InsulinYN == "Yes"]<-1
tmpDat$binary_group<-0
tmpDat$binary_group[tmpDat$Group == "Metformin"]<-1
tmpDat$binary_insulin<-as.numeric(tmpDat$binary_insulin)
tmpDat$binary_EMERGE<-0
tmpDat$binary_EMERGE[tmpDat$compositeYN == "Yes"]<-1
```

## Creating the Composites using original EMERGE data
```{r}
insulin_t_value<-t.test(tmpDat$binary_insulin[tmpDat$Group == "Metformin"], tmpDat$binary_insulin[tmpDat$Group == "Placebo"])$statistic
gw32_t_value<-t.test(tmpDat$gw32[tmpDat$Group == "Metformin"], tmpDat$gw32[tmpDat$Group == "Placebo"])$statistic
gw38_t_value<-t.test(tmpDat$gw38[tmpDat$Group == "Metformin"], tmpDat$gw38[tmpDat$Group == "Placebo"])$statistic

study_squared_composite<-(insulin_t_value^2)+(gw32_t_value^2)+(gw38_t_value^2)
study_summed_composite<-insulin_t_value+gw32_t_value+gw38_t_value

```

## Permutation Code
Permutation Function
```{r}
permutation_function<-function(x, y, z){

  permutation_summed_composite<-numeric(100000)
  permutation_squared_composite <- numeric(100000)
  i <-1

  repeat{

    tmpDat$Group<-sample(tmpDat$Group, nrow(tmpDat), replace = F)

    t1 <- t.test(x[tmpDat$Group == "Metformin"], x[tmpDat$Group == "Placebo"])$statistic
    t2 <- t.test(y[tmpDat$Group == "Metformin"], y[tmpDat$Group == "Placebo"])$statistic 
    t3 <- t.test(z[tmpDat$Group == "Metformin"], z[tmpDat$Group == "Placebo"])$statistic

    permutation_summed_composite[i] <- t1 + t2 +t3
    permutation_squared_composite[i]<-  t1^2 + t2^2 + t3^2

    i<- i +1

    if(i > 100000) break}

  list(permutation_summed_composite  = permutation_summed_composite,
    permutation_squared_composite = permutation_squared_composite)

}

```

Creating permuted null distributions using permutation function
```{r}
set.seed(24253739)
permuted_null_distributions <-permutation_function(tmpDat$gw32, tmpDat$gw38, tmpDat$binary_insulin)
permuted_null_distributions.df <-data.frame(permuted_null_distributions)
write.csv(permuted_null_distributions.df, "Permuted Null Distributions.csv", row.names = FALSE)

```

Plotting permuted null distributions
Summed test statistics composite
```{r warning=FALSE}
ggplot(permuted_null_distributions.df, aes(x= permutation_summed_composite))+
  geom_histogram(bins = 200, fill = "lightsalmon", color = "lightsalmon")+
  scale_x_continuous(limits = c(-10, 10)) +
  labs(x = "Test Statistic", y = "Frequency") +
  geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025), color = "red", linetype = "dashed", size = 0.75)+
    geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975), color = "red", linetype = "dashed", size = 0.75)+
  annotate("text", x = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025), y = 1250, label = "2.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
    annotate("text", x = quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975), y = 1250, label = "97.5th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5)+
  theme_minimal()+ theme(panel.grid.minor = element_blank()) +
  geom_vline( xintercept = study_summed_composite, color = "steelblue", size = 1 )+
  annotate("text", x = study_summed_composite, y = 1250, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5, size = 5) + theme(axis.title = element_text(size = 16))
```

Permuted Squared test statistics composite
```{r warning=FALSE}
ggplot(permuted_null_distributions.df, aes(x= permutation_squared_composite))+
  geom_histogram(bins = 200, fill = "darksalmon", color = "darksalmon")+
  scale_x_continuous(limits = c(0, 30)) +
  labs(x = "Test Statistic", y = "Frequency") +
  geom_vline(xintercept = quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95), color = "red", linetype = "dashed", size = 0.75)+
  annotate("text", x = quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95), y = 2500, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5)+
  theme_minimal()+ theme(panel.grid.minor = element_blank(),
                         axis.title = element_text(size = 16)) +
  geom_vline(xintercept = study_squared_composite, color = "steelblue", size = 1 ) +
  annotate("text", x = study_squared_composite, y = 2500, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5)
```

## Theoretical Distributions
Summed test statistics  with equal weighting
```{r}
set.seed(24253739)
multivariate <- tmpDat %>% dplyr::select(gw32, gw38, binary_insulin)
sigma <- cor(multivariate, use = "complete.obs")
w <- c(1, 1, 1)
variance <-w%*%sigma%*%w
size <- 1000000
null_normal_distribution <- rnorm(size, mean = 0, sd = sqrt(variance))
p_value <- sum(null_normal_distribution <= study_summed_composite)/size
print(paste("P-value: ", p_value))

null_normal_distribution.df <- data.frame(null_normal_distribution)
write.csv(null_normal_distribution.df, "Theoretical Summed null distribution.csv", row.names = FALSE )

```

Plotting theoretical null distribution of summed test statistic composite
```{r}
ggplot(data = null_normal_distribution.df, aes(x =null_normal_distribution )) + geom_histogram(bins =250, colour = "skyblue", fill = "skyblue") + labs(title = "Normal Null Distribution", x = "", y = "Frequency") +
  theme_minimal() +   geom_vline(xintercept = quantile(null_normal_distribution, probs = c(0.025, 0.975)), color = "black", linetype = "dashed", linewidth = 0.75) + 
  geom_vline( xintercept = study_summed_composite, color = "gold", size = 1 ) + 
  annotate("text", x = study_summed_composite, y = 10000, label = "EMERGE Composite", color = "gold", angle = 90, vjust = -0.5)
```

Squared test statistics with eigenvalue weighting
```{r}
set.seed(24253739)
multivariate <- tmpDat %>% dplyr::select(gw32, gw38, binary_insulin)
multivariate_cm <- cor(multivariate, use = "complete.obs")
eigen_values <- eigen(multivariate_cm)$values
size <- 1000000
simulated_chi_sq_matrix <- matrix(rchisq(size * 3, df = 1), ncol = 3)
eigen_values_chi_sq<- simulated_chi_sq_matrix %*%eigen_values  
theoretical_null_squared <-data.frame(eigen_values_chi_sq)

p_value <-sum(theoretical_null_squared >= study_squared_composite)/size
print(paste("P-value: ", p_value))

theoretical_null_squared <- data.frame(chi_sq = eigen_values_chi_sq)
write.csv(theoretical_null_squared, "Theoretical Squared null distribution.csv", row.names = FALSE )

```
Plotting theoretical null distribution of squared test statistic composite with eigenvalue weights
```{r}
ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(title = "", x = "", y = "") +
  theme_minimal() + 
  geom_vline(xintercept = quantile(theoretical_null_squared$chi_sq, probs = c(0.95)), color = "red", linetype = "dashed", linewidth = 0.75) +
  annotate("text", x = 8.2, y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
  theme(panel.grid.minor = element_blank())+
  geom_vline( xintercept = study_squared_composite, color = "steelblue", size = 1 ) + 
  annotate("text", x = study_squared_composite, y = 25000, label = "Composite", color = "steelblue", angle = 90, vjust = -0.5, size = 5)

```

## Latent Variable Analysis McMenamin et al.
```{r}

```

## Power Estimation
Power estimation function
```{r}
power_function <- function(x, y, z, e, n) {
  summed_TS_power_result <- numeric(1000)  
  squared_TS_power_result <- numeric(1000)  
  EMERGE_composite_power <- numeric(1000)
  i <- 1
  repeat {
    placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
    placebo_group <- placebo_group[complete.cases(placebo_group[, c(x, y, z)]), ]
    sampled_placebo <- placebo_group[sample(nrow(placebo_group), n/2, replace = TRUE), ]

    metformin_group <- tmpDat[tmpDat$Group == "Metformin", ]
    metformin_group <- metformin_group[complete.cases(metformin_group[, c(x, y, z)]), ]
    sampled_metformin <- metformin_group[sample(nrow(metformin_group), n/2, replace = TRUE), ]

    t_test_result1<- t.test(sampled_metformin[x], sampled_placebo[x])
    t_test_result2<- t.test(sampled_metformin[y], sampled_placebo[y])
    t_test_result3<- t.test(sampled_metformin[z], sampled_placebo[z])
    t_test_result4<- t.test(sampled_metformin[e], sampled_placebo[e])

    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
    squared_TS_power_result[i]<-(((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
    EMERGE_composite_power[i] <- t_test_result4$p.value
    i<- i +1
    if(i > 1000) break}
  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 1000,## permutation test summed
         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 1000, ##permutation test squared
         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 1000, ##asymptotic chi-square
         sum((EMERGE_composite_power< 0.05))/ 1000, ## EMERGE
         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 1000)))
}
```

Applying power estimation function
```{r}
sample_vector <- seq(from = 30, to = 540, by = 10)
unified_power_results <- sapply(sample_vector, function(n_value) {
  power_function("gw32", "gw38", "binary_insulin", "binary_EMERGE", n_value)
})
unified_power_results_df <-unified_power_results
unified_power_results_df <- as.data.frame(t(unified_power_results)) 
colnames(unified_power_results_df) <- c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS")
unified_power_results_df$sample_size <- sample_vector

```

Plotting power estimation
```{r}
power_100_latvar<-0.99
power_150_latvar <- 0.99

LVF_power_df <- data.frame( power = c( 0.99, 0.99),
                                   sample_size = c(100, 150),
                                   label = "Latent Variable")
power_long <- unified_power_results_df %>%
  pivot_longer(
    cols = c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS"),
    names_to = "composite_type",
    values_to = "power"
  )
write.csv(power_long, "power_long.csv", row.names = FALSE)

ggplot(power_long, aes(x = sample_size, y = power, color = composite_type)) +
  geom_line(size = 1, alpha = 0.5) +
  labs(
    title = expression(bold("")),
    x = "Sample Size",
    y = "Power"
  ) +
  scale_x_continuous(limits = c(30, NA), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 1.1)) +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 24),
    legend.title = element_blank(),
    legend.text = element_text(size = 9, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_color_manual(
    values = c(
      "Permuted: Summed TS" = "mediumseagreen",
      "Permuted: Squared TS" = "blue",
      "Asymptotic Chi-square: Eigenvalue weights TS" = "orange2",
      "EMERGE Composite" = "magenta",
      "Asymptotic Normal: Equal weights TS" = "firebrick"
    )
  ) + 
  guides(color = guide_legend(nrow = 3)) + 
  geom_point(data = LVF_power_df, aes(x = sample_size, y = power),
    shape = 21,
    fill = "black",
    color = "black",
    size = 2,
    stroke = 1.2,
    inherit.aes = FALSE,
    vjust = -1
  ) + 
  geom_text(data = LVF_power_df, aes(x = sample_size, y = power, label = label),
    vjust = -1,   
    size = 2,
    fontface = "bold",
    color = "darkred",
    inherit.aes = FALSE
  )
```

## Type I error rate Estimation
```{r}
set.seed(24253739)
type_I_error_function <- function(x, y, z, e, n) {
  summed_TS_power_result <- numeric(10000)  
  squared_TS_power_result <- numeric(10000)  
  EMERGE_composite_power <- numeric(10000)
  i <- 1
  repeat {
    sampled_data <- tmpDat[sample(nrow(tmpDat), n, replace = TRUE), ]
    sampled_data$binary_g <- rep(c(0, 1), each = n / 2) ##binary_g refers to treatment group
    sampled_metformin <- sampled_data[sampled_data$binary_g == 1,]
    sampled_placebo <- sampled_data[sampled_data$binary_g == 0,]

    t_test_result1<- t.test(sampled_metformin[x], sampled_placebo[x])
    t_test_result2<- t.test(sampled_metformin[y], sampled_placebo[y])
    t_test_result3<- t.test(sampled_metformin[z], sampled_placebo[z])
    t_test_result4<- t.test(sampled_metformin[e], sampled_placebo[e])

    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
    squared_TS_power_result[i] <- (((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
    EMERGE_composite_power[i] <- t_test_result4$p.value
    i<- i +1
    if(i > 10000) break}
  return(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025) | summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 10000, ##permuted summed composite
         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 10000, ##permutetd squared composite
         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 10000, ##theoretical squared composite
         sum((EMERGE_composite_power< 0.05))/ 10000, ##EMERGE composite
         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 10000))
}
```

Applying power estimation code and plotting
```{r}
sample_vector <- seq(from = 30, to = 540, by = 10)
type_1_error_rate <- sapply(sample_vector, function(n_value) {
  type_I_error_function("gw32", "gw38", "binary_insulin", "binary_EMERGE", n_value)
})
type_1_error_rate_df <- as.data.frame(t(type_1_error_rate)) 
colnames(type_1_error_rate_df) <- c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS")
type_1_error_rate_df$sample_size <- sample_vector

LVF_type_1_error_rate_100 <- 0.93
LVF_type_1_error_rate_200 <- 0.96
LVF_type_1_error_df <- data.frame( error_rate = c( 0.93, 0.96),
                                   sample_size = c(100, 200),
                                   label = "Latent Variable")

error_long <- type_1_error_rate_df %>%
  pivot_longer(
    cols = c("Permuted: Summed TS", "Permuted: Squared TS", "Asymptotic Chi-square: Eigenvalue weights TS", "EMERGE Composite", "Asymptotic Normal: Equal weights TS"),
    names_to = "composite_type",
    values_to = "error_rate"
  )
write.csv(error_long, "Type I error estimation.csv", row.names = FALSE)

ggplot(error_long, aes(x = sample_size, y = error_rate, color = composite_type)) +
  geom_line(size = 1, alpha = 0.5) +
  labs(
    title = expression(bold("")),
    x = "Sample Size",
    y = "Type 1 error"
  ) +
  scale_x_continuous(limits = c(30, NA), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 0.1)) +
  theme_minimal() +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 28),
    legend.title = element_blank(),
    legend.text = element_text(size = 11, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
scale_color_manual(
    values = c(
      "Permuted: Summed TS" = "mediumseagreen",
      "Permuted: Squared TS" = "blue",
      "Asymptotic Chi-square: Eigenvalue weights TS" = "orange2",
      "EMERGE Composite" = "magenta",
      "Asymptotic Normal: Equal weights TS" = "firebrick"
    )
  ) + 
  guides(color = guide_legend(nrow = 3)) +
  geom_point(data = LVF_type_1_error_df, aes(x = sample_size, y = error_rate),
    shape = 21,
    fill = "black",
    color = "black",
    size = 2,
    stroke = 1.2,
    inherit.aes = FALSE,
    vjust = -1
  ) + 
  geom_text(data = LVF_type_1_error_df, aes(x = sample_size, y = error_rate, label = label),
    vjust = -1,   
    size = 3.5,
    fontface = "bold",
    color = "darkred",
    inherit.aes = FALSE
  )

```

## Simulating toy data set

pre-processing of tmpDat and acquiring parameters required for simulating data
```{r}
tmpDat <- tmpDat[(!is.na(tmpDat$InsulinYN) & !is.na(tmpDat$gw32) & !is.na(tmpDat$gw38)), ]
met_InsulinY_data <- filter(tmpDat, binary_insulin == 1, Group == "Metformin") 
pla_InsulinY_data <- filter(tmpDat, binary_insulin == 1, Group == "Placebo")

met_InsulinN_data <- filter(tmpDat, binary_insulin == 0, Group == "Metformin")
pla_InsulinN_data <- filter(tmpDat, binary_insulin == 0, Group == "Placebo")

met_correlation_matrix_InsulinY <- cor(x=dplyr::select(met_InsulinY_data, gw32, gw38), use = "pairwise.complete.obs")
pla_correlation_matrix_InsulinY <- cor(x=dplyr::select(pla_InsulinY_data, gw32, gw38), use = "pairwise.complete.obs")

met_correlation_matrix_InsulinN <- cor(x=dplyr::select(met_InsulinN_data, gw32, gw38), use = "pairwise.complete.obs")
pla_correlation_matrix_InsulinN <- cor(x=dplyr::select(pla_InsulinN_data, gw32, gw38), use = "pairwise.complete.obs")

avg_Met_gw32_InsulinY<-mean(tmpDat$gw32[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 1], na.rm=T) ##Average gw32 for metformin group with Insulin
avg_Met_gw38_InsulinY<-mean(tmpDat$gw38[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 1], na.rm=T)## Same again for gw38

avg_Met_gw32_InsulinN<-mean(tmpDat$gw32[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 0], na.rm=T)##Average gw38 for metformin with no Insulin
avg_Met_gw38_InsulinN<-mean(tmpDat$gw38[tmpDat$Group == "Metformin" & tmpDat$binary_insulin == 0], na.rm=T)

##Now for Placebo
avg_Plcb_gw32_InsulinY<-mean(tmpDat$gw32[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 1], na.rm=T) 
avg_Plcb_gw38_InsulinY<-mean(tmpDat$gw38[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 1], na.rm=T)

avg_Plcb_gw32_InsulinN<-mean(tmpDat$gw32[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 0], na.rm=T)
avg_Plcb_gw38_InsulinN<-mean(tmpDat$gw38[tmpDat$Group == "Placebo" & tmpDat$binary_insulin == 0], na.rm=T)
```

Generating the toy data set (aka dummy data)
```{r}
fake_met_insulinY <- mvrnorm(nrow(met_InsulinY_data), c( avg_Met_gw32_InsulinY, avg_Met_gw38_InsulinY), met_correlation_matrix_InsulinY)
fake_met_insulinY <- data.frame(fake_met_insulinY)
fake_met_insulinY$Group <- "Metformin"

fake_pcb_insulinY <- mvrnorm(nrow(pla_InsulinY_data), c( avg_Plcb_gw32_InsulinY, avg_Plcb_gw38_InsulinY), pla_correlation_matrix_InsulinY)
fake_pcb_insulinY <- data.frame(fake_pcb_insulinY)
fake_pcb_insulinY$Group <- "Placebo"

fake_met_insulinN <- mvrnorm(nrow(met_InsulinN_data), c( avg_Met_gw32_InsulinN, avg_Met_gw38_InsulinN), met_correlation_matrix_InsulinN)
fake_met_insulinN <- data.frame(fake_met_insulinN)
fake_met_insulinN$Group <- "Metformin"

fake_pcb_insulinN <- mvrnorm(nrow(pla_InsulinN_data), c( avg_Plcb_gw32_InsulinN, avg_Plcb_gw38_InsulinN), pla_correlation_matrix_InsulinN)
fake_pcb_insulinN <- data.frame(fake_pcb_insulinN)
fake_pcb_insulinN$Group <- "Placebo"

fake_InsulinY_df <- data.frame(rbind(fake_met_insulinY, fake_pcb_insulinY))
fake_InsulinY_df$InsulinYN <- "Yes"
fake_InsulinY_df$ID<- c(1:(nrow(met_InsulinY_data) + nrow(pla_InsulinY_data)))
fake_InsulinY_df <- fake_InsulinY_df[, c("ID", "Group", "InsulinYN", "gw32", "gw38")]

fake_InsulinN_df <- data.frame(rbind(fake_met_insulinN, fake_pcb_insulinN))
fake_InsulinN_df$InsulinYN <- "No"
fake_InsulinN_df$ID<- c((nrow(met_InsulinY_data) + nrow(pla_InsulinY_data) + 1): nrow(tmpDat))
fake_InsulinN_df <- fake_InsulinN_df[, c("ID", "Group", "InsulinYN", "gw32", "gw38")]

dummy_data<- rbind(fake_InsulinY_df, fake_InsulinN_df)

dummy_data$binary_insulin<-0
dummy_data$binary_insulin[dummy_data$InsulinYN == "Yes"]<-1

dummy_data$binary_group<-0
dummy_data$binary_group[dummy_data$Group == "Metformin"]<-1

write.csv(dummy_data, "Simulated toy dataset.csv", row.names = FALSE)
```

table of parameters for dummy data
```{r}
dummy_data_parameters <- data.frame(group = c("Metformin, InsulinY", "Metformin, InsulinN",
                                                 "Placebo, InsulinY", "Placebo, InsulinN"),
                                       n = c(nrow(met_InsulinY_data), nrow(met_InsulinN_data),
                                                  nrow(pla_InsulinY_data), nrow(pla_InsulinN_data)),
                                       mean_gw32 = c(avg_Met_gw32_InsulinY, avg_Met_gw32_InsulinN,
                                                     avg_Plcb_gw32_InsulinY, avg_Plcb_gw32_InsulinN),
                                       mean_gw38 = c(avg_Met_gw38_InsulinY, avg_Met_gw38_InsulinN,
                                                     avg_Plcb_gw38_InsulinY, avg_Plcb_gw38_InsulinN),
                                       correlation = c(met_correlation_matrix_InsulinY[1,2], met_correlation_matrix_InsulinN[1,2],
                                                       pla_correlation_matrix_InsulinY[1,2], pla_correlation_matrix_InsulinN[1,2]))
colnames(dummy_data_parameters) <- c("Group", "N", "Mean GW32", "Mean GW38", "Correlation")
dummy_data_parameters %>%
  kable(
    caption    = "Parameters of Simulated Toy Data",
    booktabs   = TRUE,
    digits     = c(0, 0, 2, 2, 2),
    align      = c("l", "r", "r", "r", "r")
  ) %>%
  kable_styling(
    latex_options = c("striped", "hold_position"), 
    full_width    = FALSE
  )
```

Toy data adapted for McMenamins app
Require baseline gw32 and gw38 values, which we set as 0 for all except two individuals, one of which has 0.1 as the baseline value for gw32 or gw38
Additionally, the values had to be "inverted" for the mcmenamin app to work properly. This was assessed by matching the risk difference calculated by the app to that reported in the EMERGE paper.
```{r}
simulated_data <- dummy_data %>% dplyr::select(patient_id = ID, treat = binary_group, Y1 = gw32, Y2 = gw38, Ybin = binary_insulin)
simulated_data$Y1_0 <- 0
simulated_data$Y2_0 <- 0
simulated_data$Y1_0[1] <- -0.1
simulated_data$Y2_0[2] <- -0.1
simulated_data$Ybin <- ifelse(simulated_data$Ybin == 1, 0, 1)
simulated_data$Y1 <- simulated_data$Y1 * -1
simulated_data$Y2 <- simulated_data$Y2 * -1
write.csv(simulated_data, "Simulated Inverted Emerge Data.csv", row.names = FALSE)

```